Batched Bayesian Optimization via Multi-objective Acquisition Function Ensemble for Automated Analog Circuit Design

### 2018-01-29 paper/background.tex
Update
Paper update

@@

\subsection{Gaussian Process}

Basic introduction to GP

\subsection{Acquisition Function}

Introduce LCB, EI and PI

>>>>

\textcolor{red}{TODO: See how other people describe GP and BO}

@@

\textcolor{red}{TODO: The circuit performance comes from \bf{circuit} simulation}
\textcolor{red}{TODO: format, citation of HSPICE/spectre}

We handle the scenarios that topology of the analog circuit is fixed, this is
practical, as for each task, thare are usually a lot of classical topologies
that can be used. Once the topology is fixed, the designer has to choose the
optimal design parameters according to the specifications and the circuit
device model. What we want to do is to automatically search for the optimal
design parameters, this task can be formulated as a bound-constrained black-box optimization problem:

\begin{equation}
    \label{eq:MOFormulation}
    \text{minimize}~\mathrm{FOM}(\mathbf{x})
\end{equation}
where $\mathbf{x} \in \textrm{D} \subset \textrm{R}^d$ is the design variables
and $f_i(\mathbf{x})$ represents the $i$-th cared circuit performance like gain
or phase margin of an amplifier. and the $\mathrm{FOM}(\mathbf{x})$ is the
design objective. Given the design parameters $\bf{x}$, the FOM value can be
obtained via circuit simulation using softwares like HSPICE or spectre.

\subsection{Gaussian Process Regression}

\textcolor{red}{ALERT: COPIED FROM MY PAPER SUMBITTED TO DAC}

The advantage of GP is that it not only provides a prediction like other
regression models but also gives a well-calibrated uncertainty estimation. The
uncertainties would be small in the regions with densely sampled data and large
in the regions with few samples.

GP is fully characterized by a mean function $m(\mathbf{x})$ and a covariance
function $k(\mathbf{x}, \mathbf{y})$. In this work, we use the constant mean function
$m(\mathbf{x}) = \mu_0$ and the squared exponential covariance function as follows to characterize the Gaussian process.
\begin{equation}
    \label{eq:GaussianCovarianceFunction}
    k(\mathbf{x}_i, \mathbf{x}_j) = \sigma_f^2 \exp\Big(-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^T\Lambda^{-1}(\mathbf{x}_i - \mathbf{x}_j)\Big),
\end{equation}
where $\Lambda = \mathrm{diag}(l_1, \dots, l_d)$ is a diagonal matrix and $l_i$ denotes the length scale of the $i$-th dimension. $\mu_0$, $\sigma_f$ and $\Lambda$ are the hyper-parameters.

GP can be used to model a black-box function $f(x)$. Denote the hyper-parameters as a vector $\mathbf{\theta}$, given the training set
$\{X, \mathbf{y}\}$, where $X = \{\mathbf{x_1}, \dots, \mathbf{x}_N\}$, $\mathbf{y} =
(f(\mathbf{x}_1), \dots, f(\mathbf{x}_N))^T$. The log likelihood of the training
set~\cite{GPML} can be expressed as
\begin{equation}
    \log \mathrm{p}(\mathbf{y} | X, \mathbf{\theta}) \propto -\frac{1}{2}(\mathbf{y} - \mu_0)^T K_{\mathbf{\theta}}^{-1}(\mathbf{y} - \mu_0) - \frac{1}{2} \log |K_{\mathbf{\theta}}|
    \label{eq:LikeliHood}
\end{equation}
where $K_\theta({i, j}) = k(\mathbf{x}_i, \mathbf{x}_j)$. By maximizing the likelihood
function \ref{eq:LikeliHood}, we can get the maximum likelihood estimation
(MLE) of the hyper-parameters.

Given a new data point $\mathbf{x}$, the prediction of $f(\mathbf{x})$ is
not a scalar value, but a Gaussian distribution 
\begin{equation}
f(\mathbf{x}) \sim N(\mu(\mathbf{x}),
\sigma^2(\mathbf{x}))
\label{eq:GPRPred}
\end{equation}
where $\mu(\mathbf{x})$ and $\sigma^2(\mathbf{x})$ can be expressed as
\begin{equation}
    \left\{
        \begin{array}{lll}
            \mu(\mathbf{x}) &=& \mu_0 + k(\mathbf{x},X)K_{\mathbf{\theta}}^{-1}(\mathbf{y} - \mu_0) \\
            \sigma^2(\mathbf{x}) &=& \sigma_f^2 - k(\mathbf{x}, X)K_{\mathbf{\theta}}^{-1}k(X, \mathbf{x})
        \end{array}
    \right..
    \label{eq:GPRPredEqNoisy}
\end{equation}
where $k(\mathbf{x}, X) = (k(\mathbf{x}, \mathbf{x}_1), \dots, k(\mathbf{x}, \mathbf{x}_N))^T$ and
$k(X, \mathbf{x}) = k(\mathbf{x}, X)^T$.  The $\mu(\mathbf{x})$ can be viewed as the
prediction of the function value, while the $\sigma^2(\mathbf{x})$ is a measure of
uncertainty. We refer readers
to~\cite{GPML} for more details of GP.

\subsection{Bayesian Optimization}

Bayesian optimization~\cite{shahriari2016taking} was proposed for the optimization
of expensive black-box functions. It consists of two essential ingredients,
i.e., the probabilistic surrogate models and an acquisition function. The
probabilistic surrogate models provide the prediction with uncertainties. They are
refined incrementally with newly observed data. Acquisition function is used to
explore the state space based on the surrogate model optimally.

In this subsection, we introduce the single-objective Bayesian optimization,
our extension to multi-objective optimization would be described in the next
subsection.

Without loss of generality, we only consider the minimization problem. For a
black-box function $f(\mathbf{x})$, we aim to find the global minimum.
\begin{equation}
    \label{eq:BO_Formulation}
    \mathbf{x}_* = \arg\min_{x} f(\mathbf{x})
\end{equation}

The GP model is the \emph{de facto} surrogate model for Bayesian optimization.
In Bayesian optimization, the black-box function $f(x)$ is modeled by the GP
model (\ref{eq:GPRPred}). Through an acquisition function, the prediction and
its uncertainty estimation can be combined to guide the optimization
procedure efficiently.

In Bayesian optimization, the acquisition function is constructed to balance
the exploration and exploitation. By optimizing the acquisition function, the
next data point which can efficiently explore the state space.  We consider an
acquisition function lower confidence bound (LCB)~\cite{liu2014gaspad} defined
as follows.
\begin{equation}
    \label{eq:LCB}
    \mathrm{LCB}(\mathbf{x}) = \mu(\mathbf{x}) - \kappa \sigma(\mathbf{x})
\end{equation}
where $\mu(\mathbf{x})$ and $\sigma(\mathbf{x})$ are derived from the GP prediction, and
$\kappa$ is a user-defined parameter, We set $\kappa = 3$ in this work. As can
be seen, the prediction and uncertainty estimation are combined in the LCB function.

The LCB function would have a low value if $\mu(\mathbf{x})$ is low, or the
uncertainty $\sigma(\mathbf{x})$ is high. It means that we tend to visit the data
point where $f(x)$ is minimized based on the prediction and also the region
where the prediction uncertainty is large.

Starting from a set of initial data points, in each iteration of Bayesian
optimization, a GP model is trained using the existing data points. The
acquisition function, e.g., the LCB function is then constructed, and existing
optimization algorithms can be applied to optimize the acquisition function.
The optimum of the acquisition function is then selected as the next data point
to be evaluated. Note that, during the optimization of the acquisition
function, only the evaluations of the GP models are involved, which are
computationally efficient.

Besides LCB, there also exist other acquisition functions like expected
improvement~\cite{bull2011convergence}, Thompson
sampling~\cite{chapelle2011empirical}, and predictive entropy
search~\cite{hernandez2014predictive}. A portfolio of several different
acquisition functions is also possible~\cite{hoffman2011portfolio}.

### 2018-01-29 paper/intro.tex
Paper update

Automated analog IC design is very important

Traditional methods using offline model and simulated based methods are not good

Bayesian optimization has been applied, and outperformed many other algorithm

Bayesian optimization is a sequential algorithm, there is a need to parallelize it

Review current methods

Beiefly introduce my algorithm

>>>>

% % TODO: more introduction to the importance of anlog IC sizing, as the reviewers may not have much knowledge of circuit design
Automated analog circuit design has been a challenging problem for the
community of electronic design automation (EDA). Unlike digital circuits, where
the design flow is highly automated, analog circuit design still
heavily relies on designer's experience. The design parameters of analog circuits like
transistor widths and lengths need to be manually calculated based on the
specifications and the designers' understanding of the circuit. However, due to
the ever-scaling IC manufacture technology and the increasing demands for
high-performance, low-power circuits, it is getting much more difficult to meet
the performance and time-to-market requirement with manual circuit design.
Automated analog circuit design has thus attracted more research interests in
the past decade~\cite{rutenbar2007hierarchical}.

% TODO: Traditional methods using offline model and simulated based methods are
% not good
The analog circuit design automation problems can be formulated as optimization
problems, the aim is to find the optimal vector of design parameters that give
best figure of merit (FOM). Prior works about analog circuit optimization
include offline model-based approaches
~\cite{colleran2003optimization,daems2003simulation,wang2014enabling} and
simulation based approaches. The offline model-based methods try to build
accurate global model and then apply global optimization algorithms to the
cheap-to-evaluate models, the model either comes from designers' manual
derivation or from regression models like SVM and ANN. The problem of this
approach is that the accurate models are usually hard to get, for example,
in~\cite{wang2014enabling}, 100,000 randomly simulated points are used to train
a sparse polynomial model for a circuit with ten design parameters.
Simulation-based methods, instead, treat the circuits as black-box functions
that get the objective value from circuit simulations, global optimization
algorithms are directly applied to the black-box functions. For
simulation-based circuit optimization methods, meta-heuristic algorithms like
particle swarm intelligence (PSO)~\cite{phelps2000anaconda} and differential
evolution algorithm (DE)~\cite{liu2009analog} are widely used, although these
algorithm are able to explore the whole design space, they have relatively low
convergence rate and consume many circuit simulations. When the circuit
simulation takes a long time, both model-based and simulation-based approaches
can be prohibitive.

% TODO: Bayesian optimization is a sequential algorithm, there is a need to
% parallelize it
%
% TODO: cite the two 3D-IC papers of TVLSI
To reduce the circuit simulations needed by the optimization, the Gaussian
process (GP)~\cite{GPML} model has been introduced as online surrogate model to
assist the optimization. In~\cite{liu2014gaspad}, GP is combined with
differential evolution algorithm, in~\cite{lyu2017efficient}. Recently, the
Bayesian optimization (BO)~\cite{shahriari2016taking} algorithm has been applied for analog circuit
optimization, the Bayesian optimization algorithm is firstly introduced for the
automated design of general analog circuits, and has shown very good
performance compared to other simulation-based approaches,
in~\cite{wang2017efficient}, the Bayesian optimization algorithm is combined
with adaptive Monte-Carlo sampling to optimize the yield of analog circuits and
static random-access memory (SRAM).

The Bayesian optimization algorithm is a well-studied algorithm, and has
demonstrated to be a promising algorithm for automated analog circuit design,
however, the standard Bayesian optimization algorithm is sequential, it chooses
the next evaluation point by optimizing the specified acquisition fucntion,
which makes the parallelization of Bayesian optimization non-trivial. The
sequential property of BO limits its further application, as multi-core
processors are usually availble for modern computers.

% TODO: Review current methods
% TODO: See how other papers review existing batched BO methods
Batched Bayesian optimization has been a \textcolor{red}{hot} topic in the BO
literature. Related works include the \emph{simulation matching}
method~\cite{azimi2010batch}, the \emph{BUCB} (BLCB for minimization)
method~\cite{desautels2014parallelizing}, the \emph{GP-UCB-PE}
method~\cite{contal2013parallel}, the \emph{local penalization}
method~\cite{gonzalez2016batch}, the \emph{parallel predictive entropy
search}~\cite{shah2015parallel} and the \emph{qKG}
method~\cite{wu2016parallel}.

% BUCB, UCB-PE: regret bound
% PPES: entropy search
% qKG, qEI: optimal decision if multiple points are selected
% LP: very good heuristic

% TODO: Beiefly introduce my algorithm
All the above mentioned algorithms choose to use one acquisition function, and
except for the smulation mathcing method and local penalization method which
can select arbitrary acquisition function, other parallelization methods rely on
specific acquisition function, the UCB acquisition fucntion must be used for
BUCB and GP-UCB-PE, and the knowledge gradient method must be used for qKG
algorithm. As has been stated in~\cite{hoffman2011portfolio}, no one
acquisition fucntion can always outperform other acquisition functions, so
relying on one acquisition fucntion may resulting in poor performance. Also, a
portfolio strategy that ensembles multiple acquisition functions would be
helpful.

In this paper, we propose to parallize the Bayesian optimization algorithm via
a multi-objective ensemble of acquisition functions. Firstly, in each
iteration, after the GP model is updated, multiple acquisition functions are
selected, we then perform multi-objective optimization to find the \emph{Pareto
front} (PF) of the acquisition functions. The PF represents the best trade-off
between these acquisition functions, when used in sequential mode, the propose
multi-objective acquisition ensemble (MACE) strategy can be seen as a portfolio
strategy, when batched evaluation is possible, we can sample multiple points on
the PF as there are usually much more points on the PF than the batch size.

We tested the MACE algorithm using several analytical benchmark functions and
two real-world analog circuits, including an operational amplifier with
\textcolor{red}{14} design parameters and a class-E power amplifier with
\textcolor{red}{ten} design parameters. The BLCB method, local penalization
method with expected improvement acquisition fucntion (EI-LP), the qEI methods
and qEI methods are compared with MACE. The proposed MACE method achieved
competitive performance when compared with the listed state-of-the-art
algorithms.

### 2018-01-29 paper/mace.tex
Paper update

\subsection{Multi-objective optimization}

@@

>>>>

\subsection{Multi-objective Optimization}\label{sec:MOForumlation}

\textcolor{red}{See how other MOBO paper formulate multi-objective optimization}

\textcolor{red}{ALERT: COPIED FROM MY TCAS-I paper}

When we have more than one objective to optimize, the problem is formulated as:
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\mathbf{x}),~\dots~,f_m(\mathbf{x})
    \end{aligned}
\end{equation}

Unlike the single-objective problem, there is usually no \emph{best} design for
a multi-objective problem, as objectives can be conflicting, and it is unlikely
to optimize all of them simultaneously. The goal of a multi-objective
optimization algorithm is to find a set of solutions to approximate the
$\emph{Pareto front}$. A solution $\bf{x}_1$ is said to \emph{dominate} a
solution $\bf{x}_2$ if $\forall i \in \{1 \dots m\},~f_i(\bf{x}_1) \le
f_i(\bf{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bf{x}_1) < f_j(\bf{x}_2)$.
A point $\bf{x}_*$ is $\emph{weakly Pareto-optimal}$ if $\forall \bf{x} \in
\bf{\textrm{X}}$, $\bf{x}_*$ is not dominated by $\bf{x}$.  A point $\bf{x}_*$
is $\emph{Pareto-optimal}$ if $\forall x \in \bf{\textrm{X}}$, $\bf{x}_*$ is
not dominated by $\bf{x}$ and $\exists i \in \{1 \dots m\},~f_i(\bf{x}_*) <
f_i(\bf{x})$. The set of Pareto optimal outcomes is called the \emph{Pareto
front}.

@@

% TODO: discuss the difference of acquisition functions: LCB: do not avoid repeated sampling, EI: always positive
